---
title: "Test octopus API"
author: "Ben Anderson"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
---

```{r setup, include=FALSE}
# options ----
knitr::opts_chunk$set(echo = TRUE)

# libraries ----
library(data.table)
library(flextable)
library(ggplot2)
library(hms)
library(httr)
library(jsonlite)
library(lubridate)
library(skimr)

# parameters ----
source("~/octopusAPI_key.R") # api params
# DO NOT PRINT THE KEY!

rmdParams <- list() # params for this report
rmdParams$odPath <- "~/Dropbox/data/ba_octopus/" # where to save the json files
message("Test if rmdParams$odPath (", rmdParams$odPath, ") exists...")
dir.exists(path.expand(rmdParams$odPath))

# Functions ----
makeFlexTable <- function(df, cap = "caption"){
  # makes a pretty flextable - see https://cran.r-project.org/web/packages/flextable/index.html
  ft <- flextable::flextable(df)
  ft <- colformat_double(ft, digits = 1)
  ft <- fontsize(ft, size = 9)
  ft <- fontsize(ft, size = 10, part = "header")
  ft <- set_caption(ft, caption = cap)
  return(flextable::autofit(ft))
}

makeDerivedVars <- function(dt){
  # takes a consumption data.table and adds frequently used derived vars
  dt[, dv_start := lubridate::as_datetime(interval_start)] # start of half-hour
  dt[, dv_hms := hms::as_hms(dv_start)]
  dt[, dv_date := lubridate::as_date(dv_start)]
  
  dt[, dv_peakPeriod := ifelse(dv_hms < as_hms("07:00:00"), "Early morning", NA)]
  dt[, dv_peakPeriod := ifelse(dv_hms >= as_hms("07:00:00") & dv_hms < as_hms("09:00:00"), 
                               "Morning peak", 
                               dv_peakPeriod)]
  dt[, dv_peakPeriod := ifelse(dv_hms >= as_hms("09:00:00") & dv_hms < as_hms("16:00:00"), 
                               "Day time", 
                               dv_peakPeriod)]
  dt[, dv_peakPeriod := ifelse(dv_hms >= as_hms("16:00:00") & dv_hms < as_hms("20:00:00"), 
                               "Evening peak", 
                               dv_peakPeriod)]
  dt[, dv_peakPeriod := ifelse(dv_hms >= as_hms("20:00:00"), 
                               "Late evening", 
                               dv_peakPeriod)]
  
  levels <- c("Early morning", "Morning peak", "Day time", 
              "Evening peak", "Late evening") # set order nicely for plots & tables
  # check
  # dt[, dv_hour := lubridate::hour(dv_start)]
  #table(dt$dv_hour, dt$dv_peakPeriod)
  
  dt[, dv_peakPeriod := factor(dv_peakPeriod, levels)] # set order for plots etc
  
  dt[, dv_weekend := ifelse(lubridate::wday(dv_date, label = TRUE) == "Sat",
                            "Saturday",
                            "Weekday")]
  dt[, dv_weekend := ifelse(lubridate::wday(dv_date, label = TRUE) == "Sun",
                            "Sunday",
                            dv_weekend)]
  return(dt)
}
```

# Inspired by

https://www.guylipman.com/octopus/api_guide.html

and using

https://developer.octopus.energy/docs/api/#

# Test without authentication

```{r getBasic}
# test
url <- "https://api.octopus.energy/v1/products"
message("Getting: ", url)
resp <- httr::GET(url)

message("Status code: ", resp$status_code)

df <- jsonlite::parse_json(resp, simplifyVector = TRUE)
makeFlexTable(head(df$results), cap = "Example products list (first 6 rows)")
```

# Tests with authentication

## Basic info

```{r getAuth}
url <- paste0("https://api.octopus.energy/v1/accounts/", apiParams$accountNo , "/")

resp <- httr::GET(url = url, authenticate(user = apiParams$key, password = ""))

df <- jsonlite::parse_json(resp, simplifyVector = TRUE)

props <- data.table::as.data.table(df$properties)
makeFlexTable(head(props[, .(town, county, 
                             electricity_meter_points,gas_meter_points)]), cap = "Properties linked to this account (non-disclosive data)")


```

List the electricity meter points.

```{r listElec}
# this is a list of n mpans
length(props$electricity_meter_points)
message("n MPANS listed: ", length(df$properties$electricity_meter_points))
for(n in 1:length(df$properties$electricity_meter_points)){
  print(props$electricity_meter_points[n])
}

```


```{r listGas}
length(df$properties$gas_meter_points)
message("n MPRNS listed: ", length(df$properties$gas_meter_points))
for(n in 1:length(df$properties$gas_meter_points)){
  print(df$properties$gas_meter_points[n])
}
```

## Electricity

### Consumption

See: https://www.guylipman.com/octopus/api_guide.html#s3

```{r getElecConsumption}
url <- paste0("https://api.octopus.energy/v1/electricity-meter-points/", 
              apiParams$elec_import_mpan , "/",
              "meters/",
              apiParams$elec_import_serial, "/",
              "consumption/",
              "?period_from=2022-01-01T00:00Z",
              "&page_size=10000")
# get data via httr ----
resp <- httr::GET(url = url, authenticate(user = apiParams$key, password = ""))
df <- jsonlite::parse_json(resp, simplifyVector = TRUE) # creates a df of which 'results' = the data

elecCons_dt <- data.table::as.data.table(df$results) # convert to dt

# derived variables ----
elecCons_dt <- makeDerivedVars(elecCons_dt)

maxTime <- max(elecCons_dt$dv_start)

hoursAgo <- now() - maxTime
```

The data used here is up to `r maxTime`, which is `r round(hoursAgo,1)` hours ago. In general the Octopus API seems to have data up to `midnight last night`.

Figure \@ref(fig:plotElecImportHalfHourly) shows half-hourly electricity import ('consumption') for the current year. Spot the power cuts...

> To do: mark weekends somehow

```{r plotElecImportHalfHourly, fig.cap="Half hourly electricity import (current year)"}
ggplot2::ggplot(elecCons_dt, aes(x = dv_date, y = dv_hms, fill = consumption)) +
  geom_tile() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c(name = "Electricity import (kWh)") +
  labs(x = "Date",
       y = "Half-hour")
```



```{r plotElecImportDaily, fig.cap="Daily electricity import (current year)"}

plotDT <- elecCons_dt[, .(sum_kWh = sum(consumption),
                 mean_kWh = mean(consumption),
                 nObs = .N), keyby = .(dv_date, dv_weekend)]

ggplot2::ggplot(plotDT, aes(x = dv_date, y = mean_kWh, 
                            colour = dv_weekend)) +
  geom_point() +
  geom_smooth() +
  #facet_grid(dv_peakPeriod ~ .) +
  scale_colour_viridis_d(name = "Weekend") +
  labs(x = "Date",
       y = "Mean kWh")
```

Figure \@ref(fig:plotElecImportDailyByPeriod) shows the mean daily kWh import with a smoothed curve for each period as defined below.

`Early morning` is effectively our baseload.

```{r checkPeriods}
# check periods
t <- elecCons_dt[, .(min = min(dv_hms),
                max = max(dv_hms)),
            keyby = .(dv_peakPeriod)]

t
```


```{r plotElecImportDailyByPeriod, fig.cap="Daily electricity import by peak period (current year)"}
plotDT <- elecCons_dt[, .(sum_kWh = sum(consumption),
                 mean_kWh = mean(consumption),
                 nObs = .N), keyby = .(dv_date, dv_peakPeriod, dv_weekend)]

ggplot2::ggplot(plotDT, aes(x = dv_date, y = mean_kWh, 
                            colour = dv_peakPeriod)) +
  geom_line() +
  geom_smooth() +
  #facet_grid(dv_peakPeriod ~ .) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend (ncol = 3)) +
  scale_colour_viridis_d(name = "Peak period") +
  labs(x = "Date",
       y = "Mean kWh per period")
```

### Export

This will be a new MPAN but specified as export - although the url will still say `consumption`. We do not have this even though the PV is exporting on (some) days.

It may be that we only get this data if we sign up for the export tariff.

See https://www.guylipman.com/octopus/api_guide.html#s3

```{r getElecExport}

```

```{r plotElecExport}

```

## Gas Consumption

```{r getGasConsumption}
url <- paste0("https://api.octopus.energy/v1/gas-meter-points/", 
              apiParams$gas_mpan , "/",
              "meters/",
              apiParams$gas_serial, "/",
              "consumption",
              "?period_from=2022-01-01T00:00Z",
              "&page_size=10000")
resp <- httr::GET(url = url, authenticate(user = apiParams$key, password = ""))
df <- jsonlite::parse_json(resp, simplifyVector = TRUE)
gasCons_dt <- data.table::as.data.table(df$results)
gasCons_dt <- makeDerivedVars(gasCons_dt)
```

Note that this data starts later as we finally got the original un-registered smart meter replaced in February.

Figure \@ref(fig:plotGasImport) shows half-hourly gas import ('consumption') for the current year. The power cuts are even easier to see here. Interestingly the pattern after the gas boiler was serviced in June is more varied. _What did he change_?

```{r plotGasImport, fig.cap="Half-hourly gas consumption (current year)"}
ggplot2::ggplot(gasCons_dt, aes(x = dv_date, y = dv_hms, fill = consumption)) +
  geom_tile() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c(name = "Gas import (kWh)") +
  labs(x = "Date",
       y = "Half-hour")
```

Figure \@ref(fig:plotGasImportDaily) shows the mean daily kWh import with a smoothed curve.

> To do: mark weekends etc

```{r plotGasImportDaily, fig.cap="Daily gas consumption (current year)"}

plotDT <- gasCons_dt[, .(sum_kWh = sum(consumption),
                 mean_kWh = mean(consumption),
                 nObs = .N), keyby = .(dv_date, dv_weekend)]

ggplot2::ggplot(plotDT, aes(x = dv_date, y = sum_kWh, 
                            colour = dv_weekend)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend (ncol = 3)) +
  scale_colour_viridis_d(name = "Weekend") +
  labs(x = "Date",
       y = "Sum kWh per day")

ggplot2::ggplot(plotDT, aes(x = dv_date, y = mean_kWh, 
                            colour = dv_weekend)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend (ncol = 3)) +
  scale_colour_viridis_d(name = "Weekend") +
  labs(x = "Date",
       y = "Mean kWh per day")
```

Figure \@ref(fig:plotGasImportDailyByPeriod) shows the mean daily kWh import with a smoothed curve by period of the day.

> To do: mark weekends etc

```{r plotGasImportDailyByPeriod, fig.cap="Daily gas consumption by peak period (current year)"}

plotDT <- gasCons_dt[, .(sum_kWh = sum(consumption),
                 mean_kWh = mean(consumption),
                 nObs = .N), keyby = .(dv_date, dv_peakPeriod, dv_weekend)]

ggplot2::ggplot(plotDT, aes(x = dv_date, y = mean_kWh, 
                            colour = dv_peakPeriod)) +
  geom_line() +
  geom_smooth() +
  #facet_grid(dv_peakPeriod ~ .) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend (ncol = 3)) +
  scale_colour_viridis_d(name = "Peak period") +
  labs(x = "Date",
       y = "Mean kWh per period")
```

# Annexes
## Data descriptions
### Electricity consumption

Use `skmir::skim()`to summarise.

```{r skimElecCons}
skimr::skim(elecCons_dt)
```

### Gas consumption

Use `skmir::skim()`to summarise.

```{r skimGasCons}
skimr::skim(gasCons_dt)
```


# References

