---
title: "Test the octopus API and build an energy use report"
author: "Ben Anderson"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
---

```{r setup, include=FALSE}
# options ----
knitr::opts_chunk$set(echo = TRUE)

# libraries ----

library(dkUtils) # find me at https://github.com/dataknut/dkUtils

rmdLibs <- c("data.table",
             "flextable",
             "ggplot2",
             "hms",
             "httr",
             "jsonlite",
             "lubridate",
             "plotly",
             "skimr")
dkUtils::loadLibraries(rmdLibs)

# parameters ----
source("~/Dropbox/api/octopusAPI_key.R") # api params
# DO NOT PRINT THE KEY!

rmdParams <- list() # params for this report
rmdParams$odPath <- "~/Dropbox/data/ba_octopus/" # where to save the json files
message("Test if rmdParams$odPath (", rmdParams$odPath, ") exists...")
dir.exists(path.expand(rmdParams$odPath))

# Functions ----
makeFlexTable <- function(df, cap = "caption", digits = 0){
  # makes a pretty flextable - see https://cran.r-project.org/web/packages/flextable/index.html
  ft <- flextable::flextable(df)
  ft <- colformat_double(ft, digits = digits)
  ft <- fontsize(ft, size = 9)
  ft <- fontsize(ft, size = 10, part = "header")
  ft <- set_caption(ft, caption = cap)
  return(flextable::autofit(ft))
}

makeDerivedVars <- function(dt){
  # takes a consumption data.table and adds frequently used derived vars
  dt[, dv_start := lubridate::as_datetime(interval_start)] # start of half-hour
  dt[, dv_hms := hms::as_hms(dv_start)]
  dt[, dv_date := lubridate::as_date(dv_start)]
  
  dt[, dv_peakPeriod := ifelse(dv_hms < as_hms("07:00:00"), "Early morning", NA)]
  dt[, dv_peakPeriod := ifelse(dv_hms >= as_hms("07:00:00") & dv_hms < as_hms("09:00:00"), 
                               "Morning peak", 
                               dv_peakPeriod)]
  dt[, dv_peakPeriod := ifelse(dv_hms >= as_hms("09:00:00") & dv_hms < as_hms("16:00:00"), 
                               "Day time", 
                               dv_peakPeriod)]
  dt[, dv_peakPeriod := ifelse(dv_hms >= as_hms("16:00:00") & dv_hms < as_hms("20:00:00"), 
                               "Evening peak", 
                               dv_peakPeriod)]
  dt[, dv_peakPeriod := ifelse(dv_hms >= as_hms("20:00:00"), 
                               "Late evening", 
                               dv_peakPeriod)]
  
  levels <- c("Early morning", "Morning peak", "Day time", 
              "Evening peak", "Late evening") # set order nicely for plots & tables
  # check
  # dt[, dv_hour := lubridate::hour(dv_start)]
  #table(dt$dv_hour, dt$dv_peakPeriod)
  
  dt[, dv_peakPeriod := factor(dv_peakPeriod, levels)] # set order for plots etc
  
  dt[, dv_weekend := ifelse(lubridate::wday(dv_date, label = TRUE) == "Sat",
                            "Saturday",
                            "Weekday")]
  dt[, dv_weekend := ifelse(lubridate::wday(dv_date, label = TRUE) == "Sun",
                            "Sunday",
                            dv_weekend)]
  return(dt)
}
```

# Inspired by

https://www.guylipman.com/octopus/api_guide.html

and using

https://developer.octopus.energy/docs/api/#

# Code

More at: https://dataknut.github.io/octopusAPI/

Code at: https://github.com/dataknut/octopusAPI

# Test without authentication

Check the response code. This seems to generate errors (sometimes).

```{r getBasic}
# test
url <- "https://api.octopus.energy/v1/products"
message("Getting: ", url)
resp <- httr::GET(url)

message("Status code: ", resp$status_code)

# df <- jsonlite::parse_json(resp, simplifyVector = TRUE)
# makeFlexTable(head(df$results), cap = "Example products list (first 6 rows)")
```

# Tests with authentication

## Basic info

```{r getAuth}
url <- paste0("https://api.octopus.energy/v1/accounts/", apiParams$accountNo , "/")

resp <- httr::GET(url = url, authenticate(user = apiParams$key, password = ""))

df <- jsonlite::parse_json(resp, simplifyVector = TRUE)

props <- data.table::as.data.table(df$properties)
makeFlexTable(head(props[, .(town, county, 
                             electricity_meter_points,gas_meter_points)]), cap = "Properties linked to this account (non-disclosive data)")


```

List the electricity meters by MPAN. There should be only one...

```{r listElec}
# this is a list of n mpans
length(props$electricity_meter_points)
message("n MPANS listed: ", length(df$properties$electricity_meter_points))
for(n in 1:length(df$properties$electricity_meter_points)){
  print(props$electricity_meter_points[n])
}

```

List the gas meters by MPRN. There should be only one...

```{r listGas}
length(df$properties$gas_meter_points)
message("n MPRNS listed: ", length(df$properties$gas_meter_points))
for(n in 1:length(df$properties$gas_meter_points)){
  print(df$properties$gas_meter_points[n])
}
```

## Electricity

### Consumption

See: https://www.guylipman.com/octopus/api_guide.html#s3

Start data extraction from 1st Jan 2022 as we had smart meter (re)installed in January.

```{r getElecConsumption}
url <- paste0("https://api.octopus.energy/v1/electricity-meter-points/", 
              apiParams$elec_import_mpan , "/",
              "meters/",
              apiParams$elec_import_serial, "/",
              "consumption/",
              "?period_from=2022-01-01T00:00Z",
              "&page_size=10000")
# get data via httr ----
resp <- httr::GET(url = url, authenticate(user = apiParams$key, password = ""))
df <- jsonlite::parse_json(resp, simplifyVector = TRUE) # creates a df of which 'results' = the data

elecCons_dt <- data.table::as.data.table(df$results) # convert to dt

# derived variables ----
elecCons_dt <- makeDerivedVars(elecCons_dt)

maxTime <- max(elecCons_dt$dv_start)

hoursAgo <- now() - maxTime

# meter is SMETS2
elecCons_dt[, consumption_kWh := consumption] # for clarity - see https://developer.octopus.energy/docs/api/#list-consumption-for-a-meter
```

The data used here is up to `r maxTime`, which is `r round(hoursAgo,1)` hours ago. In general the Octopus API seems to have data up to `midnight last night`.

```{r elecTotal}
t <- elecCons_dt[, .(nDays = uniqueN(dv_date),
                     sumkWh = sum(consumption_kWh),
                     hh_meankWh = mean(consumption_kWh),
                     hh_minkWh = min(consumption_kWh),
                     hh_maxkWh = max(consumption_kWh)),
                 keyby = .(month = lubridate::month(dv_date))]

makeFlexTable(t, cap = "Monthly stats", digits = 2)
message("Total elec to date")
sum(t$sumkWh)

t <- elecCons_dt[, .(nDays = uniqueN(dv_date),
                     sumkWh = sum(consumption_kWh)),
                 keyby = .(dv_date)]
message("########")
elec_kWh <- 0.24
message("Projected elec annual total @ ", elec_kWh,"p")
annualTotal <- mean(t$sumkWh)*365
annualTotal
message("Projected annual elec cost @ ", elec_kWh,"p - including standing charge")
annualCost <- (annualTotal*elec_kWh)+(365*0.2401) # standing charge
annualCost
message("Projected mean monthly elec cost @ ", elec_kWh,"p - including standing charge")
annualCost/12

message("########")
elec_kWh <- 0.34
message("Projected elec annual total @ ", elec_kWh,"p")
annualTotal <- mean(t$sumkWh)*365
annualTotal
message("Projected annual elec cost @ ", elec_kWh,"p - including standing charge")
annualCost <- (annualTotal*elec_kWh)+(365*0.2401) # standing charge
annualCost
message("Projected mean monthly elec cost @ ", elec_kWh,"p - including standing charge")
annualCost/12
```

> TO DO: fix £ analysis to use tariff from API

URL will be something like

https://api.octopus.energy/v1/products/LOYAL-FIX-12M-21-10-07/electricity-tariffs/E-1R-LOYAL-FIX-12M-21-10-07-A/standard-unit-rates/

### Half-hourly analysis

Figure \@ref(fig:plotElecImportHalfHourly) shows half-hourly electricity import ('consumption') for the current year. Spot the power cuts...

> To do: mark weekends somehow

```{r plotElecImportHalfHourly, fig.cap="Half hourly electricity import (current year)"}
ggplot2::ggplot(elecCons_dt, aes(x = dv_date, y = dv_hms, fill = consumption_kWh)) +
  geom_tile() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c(name = "Electricity import (kWh)") +
  labs(x = "Date",
       y = "Half-hour")
```

Repeat but with just the last 7 days of data - useful for checking recent appliance use and offspring effects.

```{r plotElecImportHalfHourlyRecent, fig.cap="Half hourly electricity import (current year, last 7 days)"}
today <- lubridate::today()
plotDT <- elecCons_dt[dv_date >= today - 7]
p <- ggplot2::ggplot(plotDT, aes(x = dv_date, y = dv_hms, fill = consumption_kWh)) +
  geom_tile() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c(name = "Electricity import (kWh)") +
  labs(x = "Date",
       y = "Half-hour")

plotly::ggplotly(p)

plotDT[, dow := lubridate::wday(dv_date, label = TRUE)]
recent_dt <- plotDT[, .(sum_kWh_elec = sum(consumption_kWh)), keyby = .(dv_date, dow, dv_peakPeriod)]
daily_totals <- plotDT[, .(Total = sum(consumption_kWh)), keyby = .(dv_date, dow)]

t <- dcast(recent_dt, dv_date + dow ~ dv_peakPeriod, val.var = sum_kWh_elec)
# add totals
t <- t[daily_totals]
makeFlexTable(t, digits = 2,
              cap = "Recent electricity use")
```

### Daily analysis

```{r plotElecImportDaily, fig.cap="Daily electricity import (current year)"}

plotDT <- elecCons_dt[, .(sum_kWh = sum(consumption_kWh),
                 mean_kWh = mean(consumption_kWh),
                 nObs = .N), keyby = .(dv_date, dv_weekend)]

ggplot2::ggplot(plotDT, aes(x = dv_date, y = mean_kWh, 
                            colour = dv_weekend)) +
  geom_point() +
  geom_smooth() +
  #facet_grid(dv_peakPeriod ~ .) +
  scale_colour_viridis_d(name = "Weekend") +
  labs(x = "Date",
       y = "Mean kWh")
```

Figure \@ref(fig:plotElecImportDailyByPeriod) shows the mean daily kWh import with a smoothed curve for each period as defined below.

`Early morning` is effectively our baseload.

```{r checkPeriods}
# check periods
t <- elecCons_dt[, .(min = min(dv_hms),
                max = max(dv_hms)),
            keyby = .(dv_peakPeriod)]

t
```


```{r plotElecImportDailyByPeriod, fig.cap="Daily electricity import by peak period (current year)"}
plotDT <- elecCons_dt[, .(sum_kWh = sum(consumption_kWh),
                 mean_kWh = mean(consumption_kWh),
                 nObs = .N), keyby = .(dv_date, dv_peakPeriod, dv_weekend)]

ggplot2::ggplot(plotDT, aes(x = dv_date, y = mean_kWh, 
                            colour = dv_peakPeriod)) +
  geom_line() +
  geom_smooth() +
  #facet_grid(dv_peakPeriod ~ .) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend (ncol = 3)) +
  scale_colour_viridis_d(name = "Peak period") +
  labs(x = "Date",
       y = "Mean kWh per period")
```

### PV generation

Not Octopus data - irregular reads by us. Load from .xslsx

Figure \@ref(fig:getPVgen) compares our PV generation with overall grid import. These are mean values. We are occasionally exporting (Table \@ref(tab:elecTotal) has zeros) but we do not (currently) have access to the export data.

```{r getPVgen, fig.cap="Compare grid import and PV gen"}
library(openxlsx)
f <- "~/Dropbox/Home/houses/whiteHorseMews/2WhiteHorseMewsRunningCosts.xlsx"
pvGen <- openxlsx::read.xlsx(f,
                             sheet = "PV",
                             detectDates = TRUE)
pvGen_DT <- data.table::as.data.table(pvGen)
pvGen_DT[, meankWhGen := diff/n.days]
pvGen_DT[, dv_date := lubridate::as_datetime(Date)]
pvGen_DT[, month := lubridate::month(dv_date)]

monthlyPvGen_DT <- pvGen_DT[, .(hh_meankWh = mean(meankWhGen)/48 # input value is mean per day
                                   ),
                              keyby = .(month = lubridate::month(dv_date),
                                        year = lubridate::year(dv_date))
                              ]
monthlyPvGen_DT[, value := "PV generation"]

monthlyElec_DT <- elecCons_dt[, .(hh_meankWh = mean(consumption_kWh)),
                              keyby = .(month = lubridate::month(dv_date),
                                        year = lubridate::year(dv_date))]
monthlyElec_DT[, value := "Grid import"]

plotDT <- rbind(monthlyPvGen_DT, monthlyElec_DT)

ggplot2::ggplot(plotDT, aes(x = month, y = hh_meankWh, 
                            colour = value)) +
  geom_line()+
  scale_color_discrete(name = "Type")+
  facet_grid(year ~ .)
```

> To do: model value of PV gen if we were to use all of it. 

### PV Export

This will be a new MPAN but specified as export - although the url will still say `consumption`. We do not have this even though the PV is exporting on (some) days.

It may be that we only get this data if we sign up for the export tariff.

See https://www.guylipman.com/octopus/api_guide.html#s3

```{r getElecExport}

```

```{r plotElecExport}

```


### Electricity emissions

In theory our emissions from electricity use are zero because we are on a renewable-only tariff. But life is not so simple. We don't have a private wire to a wind turbine so the electrons we import (stick with it) are as averagely green as all the rest.

We also con't avoid the 'Well To Tank' emissions and those associated with transmission losses.

To further complicate things there are at least two different ways to estimate our emissions.

 * use the annual BEIS emissions factor and multiply by the kWh in question - be it half-hourly, daily, annual, whatever. That's the simple way.
 * use the NG-ESO half-hourly emissions factors which reflect the generation mix (with some caveats) of the grid at half-hourly intervals
 
`Does it matter?` you cry. Well it might. If we've been able to 'flex' our usage in line with \@[theBakingForecast](https://twitter.com/baking4cast) then who knows, maybe we'll be concentrating our usage in times when the grid is actually drawing on more renewables.

So let's take a look. We'll do both the BEIS-based and NG-ESO based calculations to see. For now we'll ignore the WTT and the T&D losses to keep the results comparable. We'll come back to that later.

```{r setBEISelec}
rmdParams$BEIS_elec_ci <-   0.21233 
```

For the BEIS method, we'll have to use the [2021 emissions factor](https://www.gov.uk/government/collections/government-conversion-factors-for-company-reporting) as the 2022 value is not yet available.

For 2021 this is: `r rmdParams$BEIS_elec_ci` Kg CO2e/kWh

For the NG-ESO method we use the NG-ESO [half-hourly carbon intensity data](https://data.nationalgrideso.com/carbon-intensity1/historic-generation-mix) that match to the half-hours in our electricity use dataset.

We save this locally, if it is older than 1 day we re-download.

```{r setNGESOelec}
# this needs to be more clever - only download dates we want from API?
esoF <- here::here("data", "latest_df_fuel_ckan.csv")
if(file.exists(paste0(esoF, ".gz"))){
  message("We already have a version saved to: ", paste0(esoF, ".gz"))
  message("Loading it...")
  ngeso_dt_orig <- data.table::fread(paste0(esoF, ".gz"))
} else {
  message("We don't already have a version, downloading and saving to: ", esoF)
  ngeso_dt_orig <- data.table::fread("https://data.nationalgrideso.com/backend/dataset/88313ae5-94e4-4ddc-a790-593554d8c6b9/resource/f93d1835-75bc-43e5-84ad-12472b180a98/download/df_fuel_ckan.csv")
  # nice dateTime
  ngeso_dt_orig[, dv_start := lubridate::as_datetime(DATETIME)]
  data.table::fwrite(ngeso_dt_orig, esoF)
  dkUtils::gzipIt(esoF)
  data.table::fwrite(ngeso_dt_orig, "~/Dropbox/data/UK_NGESO/genMix/latest_df_fuel_ckan.csv") # save locally for future re-use
  dkUtils::gzipIt("~/Dropbox/data/UK_NGESO/genMix/latest_df_fuel_ckan.csv")
}

# if older than 1 day, reload
today <- lubridate::today()
lastNGESO <- as.Date(max(ngeso_dt_orig$dv_start))

if(today - lastNGESO > 1) {
  # old data, reload
  message("But the version we have dates from ", lastNGESO, " (",today - lastNGESO ," days ago), downloading latest...")
  ngeso_dt_orig <- data.table::fread("https://data.nationalgrideso.com/backend/dataset/88313ae5-94e4-4ddc-a790-593554d8c6b9/resource/f93d1835-75bc-43e5-84ad-12472b180a98/download/df_fuel_ckan.csv")
  # nice dateTime
  ngeso_dt_orig[, dv_start := lubridate::as_datetime(DATETIME)]
  data.table::fwrite(ngeso_dt_orig, esoF)
  dkUtils::gzipIt(esoF)
}


# merge to usage data
setkey(ngeso_dt_orig, dv_start)
setkey(elecCons_dt, dv_start)
elecCons_dt <- ngeso_dt_orig[, .(dv_start, CARBON_INTENSITY, LOW_CARBON_perc, RENEWABLE_perc)][elecCons_dt] # keeps match to our electricity use

# there will be NAs if some datetimes are missing from ngeso_dt_orig

# we think renewable is wind + solar, low carbon includes nuclear
#ggplot2::ggplot(t, aes(x = RENEWABLE_perc, y = LOW_CARBON_perc)) +
#  geom_point()
```

Mean half-hourly carbon intensity from the NG-ESO data for the data period was `r round(mean(elecCons_dt$CARBON_INTENSITY, na.rm = TRUE)/1000,4)` Kg CO2e/kWh. If this is substantially different to  the BEIS 2021 value of `r round(rmdParams$BEIS_elec_ci,4)` Kg CO2e/kWh, we would  expect emissions estimates using the NG-ESO factor to differ.

For context, Figure \@ref(fig:ngesoMonthly) summarises the mean half-hourly carbon intensity by month for the data period. We can clearly see that February 2022 was a very low carbon month... in fact it was a [very windy month with 3 named storms](https://www.metoffice.gov.uk/binaries/content/assets/metofficegovuk/pdf/weather/learn-about/uk-past-events/summaries/uk_monthly_climate_summary_202202a.pdf).

```{r ngesoMonthly, fig.cap="Monthly mean carbon intensity for the data period by month (NG-ESO data)"}
elecCons_dt[, dv_month := lubridate::month(dv_date, label = TRUE)]
ggplot2::ggplot(elecCons_dt, aes(x = dv_month, y = CARBON_INTENSITY)) +
  geom_violin(draw_quantiles = c(0.5)) +
  #geom_jitter() +
  #geom_boxplot() +
  labs(x = "Month",
       y = "Half-hourly carbon intensity",
       caption = "Median drawn")

```

Figure \@ref(fig:renewableConsumption) shows our half-hourly electricity kWh use vs halfhourly carbon intensity. Ideally we want a negative correlation showing that we use the most electricity when it is 'greenest' (carbon intensity is lowest). Doesn't look too good, aye?

```{r renewableConsumption, fig.cap = "Own half-hourly electricity kWh vs NG-ESO halfhourly carbon intensity"}
ggplot2::ggplot(elecCons_dt, aes(x = CARBON_INTENSITY, y = consumption_kWh, colour = RENEWABLE_perc)) +
  geom_point() +
  facet_wrap(. ~ dv_peakPeriod) +
  geom_smooth() +
  scale_color_continuous(name = "% renewables", low = "red", high = "green") +
  theme(legend.position = "bottom") +
    labs(y = "Halfhourly electricity kWh")

```

What if we visualise using a box plot according to carbon intensity decile? So this means we divide the carbon intensity values into 10 equal groups - deciles. This is Figure \@ref(fig:halfhourlyCIbox). Doesn't look too good either - our median usage (the horizontal bar in the boxes) seems to trend slightly upwards as we move to higher carbon intensity deciles.

```{r halfhourlyCIbox, fig.cap="Boxing clever"}
# this will generate NAs if the CI data is missing for some of the (most recent) dateTimes 
elecCons_dt[, CI_deciles := cut_number(CARBON_INTENSITY, n = 10)]

ggplot2::ggplot(elecCons_dt, aes(x = CI_deciles, y = consumption_kWh)) +
  geom_boxplot() +
  labs(y = "Halfhourly electricity kWh",
       x = "Carbon intensity decile")
```

So what if we just add up all our electricity kWh by carbon intensity decile? Do we use more low carbon kWh? This is Figure \@ref(fig:consSum). Nah. The bakingforecast isn't going to like us...

```{r consSum, fig.cap="Sum of electricity kWh by carbon intensity decile"}
t <- elecCons_dt[, .(sumkWh = sum(consumption_kWh, na.rm = TRUE),
                     meankWh = mean(consumption_kWh, na.rm = TRUE)),
                 keyby = .(CI_deciles)]

ggplot2::ggplot(t, aes(x = CI_deciles, y = sumkWh)) +
  geom_col() +
  labs(y = "Sum kWh",
       x = "Carbon intensity decile")
```

Out of interest, do our emissions values look very different if we apply the BEIS 2021 annual factor to our total electricity kWh to date compared to applying the NG-ESO half-hourly values?

```{r compareMethods}
elecCons_dt[, KgCO2_ngeso := consumption_kWh * (CARBON_INTENSITY/1000)] # convert to kg
t <- elecCons_dt[, .(sumkWh = sum(consumption_kWh),
                     sumKgCO2_ngeso = sum(KgCO2_ngeso, na.rm = TRUE))]

t[, sumKgCO2_beis :=  sumkWh * rmdParams$BEIS_elec_ci]

makeFlexTable(t, cap = "Comparing emissions estimation methods using electricity kWh to date")

t <- elecCons_dt[, .(sumkWh = sum(consumption_kWh),
                     sumKgCO2_ngeso = sum(KgCO2_ngeso, na.rm = TRUE)),
                 keyby = .(dv_month)]

t[, sumKgCO2_beis :=  sumkWh * rmdParams$BEIS_elec_ci]

plotDT <- melt(t, id.vars = "dv_month")

ggplot2::ggplot(plotDT[ variable != "sumkWh",], aes(x = dv_month, y = value, fill = variable)) +
  geom_col(position = "dodge") +
  scale_color_discrete(name = "Method") +
  labs(y = "Kg CO2",
       x = "Month")


```

As we'd expect from the comparison of the values above, Table \@ref(tab:compareMethods) suggests that it does. In fact our 'in use' NG-ESO based emissions are `r round(100*(t$sumKgCO2_ngeso/t$sumKgCO2_beis),2)` % of our BEIS-based emissions depending on the month in question.

If we compare the monthly values we can see the biggest difference was in February, a month we have already identified as being more 'low carbon' (see Figure \@ref(fig:ngesoMonthly)).

## Gas Consumption

We need to convert the gas consumption from m3 to kWh - see https://developer.octopus.energy/docs/api/#list-consumption-for-a-meter

```{r set_gasM3TokWh}
gasM3TokWh <- 11.36
```

We use a multiplier of `r gasM3TokWh` kWh/m3 (https://www.theenergyshop.com/guides/how-to-convert-gas-units-to-kwh)

```{r getGasConsumption}
url <- paste0("https://api.octopus.energy/v1/gas-meter-points/", 
              apiParams$gas_mpan , "/",
              "meters/",
              apiParams$gas_serial, "/",
              "consumption",
              "?period_from=2022-01-01T00:00Z",
              "&page_size=10000")
resp <- httr::GET(url = url, authenticate(user = apiParams$key, password = ""))
df <- jsonlite::parse_json(resp, simplifyVector = TRUE)
gasCons_dt <- data.table::as.data.table(df$results)
gasCons_dt <- makeDerivedVars(gasCons_dt)

# gas 'consumption' is m3 - https://developer.octopus.energy/docs/api/#list-consumption-for-a-meter
# convert to kWh
gasCons_dt[, consumption_m3 := consumption]
gasCons_dt[, consumption_kWh := consumption * gasM3TokWh]
```

Note that this data starts later as we finally got the original un-registered smart meter replaced in February.

```{r gasTotal}


t <- gasCons_dt[, .(nDays = uniqueN(dv_date),
                     sumkWh = sum(consumption_kWh),
                     halfHourly_meankWh = mean(consumption_kWh)),
                 keyby = .(month = lubridate::month(dv_date))]

makeFlexTable(t, cap = "Monthly stats")
message("Total gas to date")
sum(t$sumkWh)

t <- gasCons_dt[, .(nDays = uniqueN(dv_date),
                     sumkWh = sum(consumption_kWh)),
                 keyby = .(dv_date)]

message("########")
gas_kWh <- 0.0719
message("Projected gas annual total @ ", gas_kWh,"p")
annualTotal <- mean(t$sumkWh)*365
annualTotal
message("Projected annual gas cost @ ", gas_kWh,"p - including standing charge")
annualCost <- (annualTotal*gas_kWh)+(365*0.261) # standing charge
annualCost
message("Projected mean monthly gas cost @ ", gas_kWh,"p - including standing charge")
annualCost/12

message("########")
gas_kWh <- 0.103
message("Projected gas annual total @ ", gas_kWh,"p")
annualTotal <- mean(t$sumkWh)*365
annualTotal
message("Projected annual gas cost @ ", gas_kWh,"p - including standing charge")
annualCost <- (annualTotal*gas_kWh)+(365*0.261) # standing charge
annualCost
message("Projected mean monthly gas cost @ ", gas_kWh,"p - including standing charge")
annualCost/12
```

### Half-hourly anlaysis

Figure \@ref(fig:plotGasImport) shows half-hourly gas import ('consumption') for the current year. The power cuts are even easier to see here. Interestingly the pattern after the gas boiler was serviced in June is more varied. _What did he change_?

```{r plotGasImport, fig.cap="Half-hourly gas consumption (current year)"}
ggplot2::ggplot(gasCons_dt, aes(x = dv_date, y = dv_hms, fill = consumption_kWh)) +
  geom_tile() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c(name = "Gas import (kWh)") +
  labs(x = "Date",
       y = "Half-hour")
```

Repeat but with just the last 7 days of data - useful for checking recent appliance use and offspring effects.

```{r plotGasImportHalfHourlyRecent, fig.cap="Half hourly gas import (current year, last 7 days)"}
today <- lubridate::today()
plotDT <- gasCons_dt[dv_date >= today - 7]
p <- ggplot2::ggplot(plotDT, aes(x = dv_date, y = dv_hms, fill = consumption_kWh)) +
  geom_tile() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c(name = "Gas import (kWh)") +
  labs(x = "Date",
       y = "Half-hour")

plotly::ggplotly(p)

plotDT[, dow := lubridate::wday(dv_date, label = TRUE)]
recent_dt <- plotDT[, .(sum_kWh_elec = sum(consumption_kWh)), keyby = .(dv_date, dow, dv_peakPeriod)]
daily_totals <- plotDT[, .(Total = sum(consumption_kWh)), keyby = .(dv_date, dow)]

t <- dcast(recent_dt, dv_date + dow ~ dv_peakPeriod, val.var = sum_kWh_elec)
# add totals
t <- t[daily_totals]
makeFlexTable(t, digits = 2,
              cap = "Recent gas use")
```

### Daily analysis

Figure \@ref(fig:plotGasImportDaily) shows the mean daily kWh import with a smoothed curve.

> To do: mark weekends etc

```{r plotGasImportDaily, fig.cap="Daily gas consumption (current year)"}

plotDT <- gasCons_dt[, .(sum_kWh = sum(consumption_kWh),
                 mean_kWh = mean(consumption_kWh),
                 nObs = .N), keyby = .(dv_date, dv_weekend)]

ggplot2::ggplot(plotDT, aes(x = dv_date, y = sum_kWh, 
                            colour = dv_weekend)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend (ncol = 3)) +
  scale_colour_viridis_d(name = "Weekend") +
  labs(x = "Date",
       y = "Sum kWh per day")
```

Repeat for mean

```{r plotGasImportDailyMean}
ggplot2::ggplot(plotDT, aes(x = dv_date, y = mean_kWh, 
                            colour = dv_weekend)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend (ncol = 3)) +
  scale_colour_viridis_d(name = "Weekend") +
  labs(x = "Date",
       y = "Mean kWh per day")
```

Figure \@ref(fig:plotGasImportDailyByPeriod) shows the mean daily kWh import with a smoothed curve by period of the day.

> To do: mark weekends etc

```{r plotGasImportDailyByPeriod, fig.cap="Daily gas consumption by peak period (current year)"}

plotDT <- gasCons_dt[, .(sum_kWh = sum(consumption_kWh),
                 mean_kWh = mean(consumption_kWh),
                 nObs = .N), keyby = .(dv_date, dv_peakPeriod, dv_weekend)]

ggplot2::ggplot(plotDT, aes(x = dv_date, y = mean_kWh, 
                            colour = dv_peakPeriod)) +
  geom_line() +
  geom_smooth() +
  #facet_grid(dv_peakPeriod ~ .) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend (ncol = 3)) +
  scale_colour_viridis_d(name = "Peak period") +
  labs(x = "Date",
       y = "Mean kWh per period")
```

### Gas emissions

This is much more simple. We can only apply the BEIS 2021 value as there are no time-varying emissions factors for gas.

```{r setBEISgas}
rmdParams$BEIS_gas_ci <- 0.20297 
```

As before, for the BEIS method we'll have to use the [2021 emissions factor](https://www.gov.uk/government/collections/government-conversion-factors-for-company-reporting) as the 2022 value is not yet available.

For 2021 this is: `r rmdParams$BEIS_gas_ci` Kg CO2e/kWh

```{r gasEmissionsTotal}
gasCons_dt[, KgCO2_beis := consumption_kWh * rmdParams$BEIS_gas_ci] 
t <- gasCons_dt[, .(sumkWh = sum(consumption_kWh),
                     sumKgCO2_beis = sum(KgCO2_beis))]

makeFlexTable(t, cap = "Emissions estimation using gas kWh to date")
```


# Annexes
## Data descriptions
### Electricity consumption

Use `skmir::skim()`to summarise.

```{r skimElecCons}
skimr::skim(elecCons_dt)
```

### Gas consumption

Use `skmir::skim()`to summarise.

```{r skimGasCons}
skimr::skim(gasCons_dt)
```


### NG-ESO data

Figure \@ref(fig:ngesoTile) shows the NG-ESO half-hourly carbon intensity over time for the data period as context. 

```{r ngesoTile, fig.cap="Half-hourly carbon intensity over time for the data period"}
ggplot2::ggplot(elecCons_dt, aes(x = dv_date, y = dv_hms, fill = CARBON_INTENSITY)) +
  geom_tile() +
  scale_fill_continuous(name = "Carbon intensity", low = "green", high = "red") +
  labs(x = "Date",
       y = "Time of day",
       caption = "Source: NG-ESO (https://data.nationalgrideso.com/carbon-intensity1/historic-generation-mix)")

```


# References

